Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-05-19 Paper/paper.tex
Update Figure 4 and Table 2

The luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.  We evaluate the double
integral in Equation~(\ref{eqn:vi}) by the Euler method, i.e., by
simply summing over the ``tiles'' of the selection function map,
without interpolating between the redshift and luminosity values of
neighbouring tiles. 
%%GW: We should add that interpolation of selection functions is not straightforward (strong gradients) and not unique. Many selection functions are very coarse. V=0 means that the selection function has systematic errors (a QSO was found where the selection probability is zero). By removing them we avoid numerical issues, but underestimate the QLF. Surveys need to report selection functions that are consistent with their actual samples.
This may result in $V_i=0$ for some quasars, in
which case we remove them from our analysis.
%%GW: Poisson statistics will underestimate the error since uncertainties in the selection function are not considered (all surveys assume a fixed selection function). For small surveys the Poisson error dominates, but for large surveys the error is driven by systematics in the selection function. These facts have been neglected so far, and we should make this point very clearly.
In each bin, we estimate the uncertainty in the luminosity function by
assuming Poisson statistics \citep{1986ApJ...303..336G} for the number
of quasars, i.e.\ assuming negligible uncertainty in the selection function.
The resultant binned luminosity function estimates are shown by the
points in Figure~\ref{fig:mosaic}.

%%GW: I've sharpened the discussion below.
As seen in Figure~\ref{fig:mosaic}, the distribution of luminosity
function values in each redshift bin are suggestive of a double power
law form for the QLF.  We will fit such a form below. 
However, in several redshift bins we note a suspicious decline of the
luminosity function at the faint limit of several surveys, for example in
the 2SLAQ sample at $z<2.2$, and in the SDSS sample at $z<1.8$ and $z\sim 4$.
The inconsistency between the SDSS faint end and the deeper
2SLAQ QLF indicates that the SDSS selection function is
systematically overestimated at its magnitude limit.
At $z\sim 4$ the SDSS faint end QLF is inconsistent with the fainter
\citet{2011ApJ...728L..26G} QLF. We identify such discrepant bins
by eye and discard them from our analysis.
%%GW: We should also set the selection functions to zero in these bins.
The discarded magnitude bins are shown in Figure~\ref{fig:mosaic} by open circles.

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.7\textwidth]{evolution_individuals.pdf}
    % summary_fromFile.py 
  \end{center}
  \caption{Redshift evolution of the four double power law parameters from the
    redshift bins shown in Figure~\ref{fig:mosaic}.  Vertical error
    bars show one-sigma (68.26\%) statistical uncertainties, whereas horizontal
    error bars show widths of the redshift bins.  We identify the
    general evolutionary trends of each of these parameters from the
    bins shown by the filled symbols.  The open symbols show bins that
    appear to be offset from these trends, likely due to unknown
    systematic errors.  The open circles at $2.2<z<3.5$ show the
    BOSS sample, while the bins at $z < 0.6$ contain AGN from the SDSS
    and 2SLAQ data sets.  See Section~\ref{sec:bins} for further
    details.}
  \label{fig:evoln}
\end{figure*}

%%GW: I would delete the z_bin column from the table. Otherwise the reader might be confused what you are plotting in the figure. The bin centers and the mean values are very similar anyway.
%%GW: Again 2.5<z<2.6 is missing
%%GW: I would add a column or a comment which redshift bins we consider reliable. I'm afraid lazy people will just take the values without thinking.
\begin{table*}
  % bins_tabulate.py; Nqso added by hand
  \caption{Best-fit double power law luminosity function parameters in
    various redshift bins.  The luminosity function parameters
    $\phi_*$, $M_*$, $\alpha$, and $\beta$ are defined in
    Equation~(\ref{eqn:dpl}), with $\beta$ denoting the faint-end
    slope.  Quasars in each bin have redshifts $z_\mathrm{min}\leq z <
    z_\mathrm{max}$.  The number of quasars in each bin is given by
    $N_\mathrm{qso}$.  The bin centre and sample mean redshift are
    given by $z_\mathrm{bin}$ and $\langle z\rangle$, respectively.
    Errors indicate one-sigma (68.26\%) uncertainties.  These values
    are shown in Figure~\ref{fig:evoln}.  The corresponding luminosity
    functions are shown in Figure~\ref{fig:mosaic}.  See
    Section~\ref{sec:bins} for further details.}
  \label{tab:bins}
  \begin{tabular}{ccccr....}
    \hline
    $\langle z\rangle$ &
    $z_\mathrm{bin}$ &
    $z_\mathrm{min}$ &
    $z_\mathrm{max}$ &
    $N_\mathrm{qso}$ &
    \multicolumn{1}{c}{$\log_{10}(\phi_*/$} &
    \multicolumn{1}{c}{$M_*$} &
    \multicolumn{1}{c}{$\alpha$} &
    \multicolumn{1}{c}{$\beta$} \\

    &
    &
    &
    &
    &
    \multicolumn{1}{c}{cMpc$^{-3}$mag$^{-1}$)} &
    &
    & \\
    \hline
    0.31 & 0.25 & 0.10 & 0.40 & 3520 & -5.62^{+0.09}_{-0.10} & -21.37^{+0.21}_{-0.22} & -2.94^{+0.09}_{-0.10} & -1.22^{+0.18}_{-0.16} \\
    0.50 & 0.50 & 0.40 & 0.60 & 4555 & -6.12^{+0.04}_{-0.04} & -23.00^{+0.08}_{-0.08} & -3.37^{+0.07}_{-0.08} & -1.36^{+0.04}_{-0.04} \\
    0.72 & 0.70 & 0.60 & 0.80 & 3308 & -6.42^{+0.09}_{-0.08} & -24.03^{+0.15}_{-0.13} & -3.55^{+0.12}_{-0.13} & -1.75^{+0.05}_{-0.04} \\
    0.91 & 0.90 & 0.80 & 1.00 & 3922 & -6.39^{+0.08}_{-0.07} & -24.51^{+0.12}_{-0.11} & -3.69^{+0.10}_{-0.11} & -1.73^{+0.06}_{-0.05} \\
    1.10 & 1.10 & 1.00 & 1.20 & 6367 & -6.53^{+0.03}_{-0.03} & -25.14^{+0.04}_{-0.04} & -4.17^{+0.08}_{-0.09} & -1.73^{+0.02}_{-0.02} \\
    1.30 & 1.30 & 1.20 & 1.40 & 6479 & -6.43^{+0.04}_{-0.04} & -25.29^{+0.06}_{-0.06} & -3.97^{+0.08}_{-0.08} & -1.74^{+0.03}_{-0.03} \\
    1.50 & 1.50 & 1.40 & 1.60 & 7539 & -6.51^{+0.03}_{-0.03} & -25.68^{+0.04}_{-0.04} & -4.28^{+0.08}_{-0.08} & -1.75^{+0.02}_{-0.02} \\
    1.71 & 1.70 & 1.60 & 1.80 & 6788 & -6.28^{+0.04}_{-0.04} & -25.55^{+0.05}_{-0.05} & -3.96^{+0.07}_{-0.07} & -1.60^{+0.03}_{-0.03} \\
    1.98 & 2.00 & 1.80 & 2.20 & 10478 & -6.72^{+0.03}_{-0.03} & -26.26^{+0.04}_{-0.04} & -4.20^{+0.07}_{-0.07} & -1.87^{+0.02}_{-0.02} \\
    2.25 & 2.25 & 2.20 & 2.30 & 4102 & -5.98^{+0.10}_{-0.09} & -25.26^{+0.19}_{-0.18} & -3.21^{+0.15}_{-0.16} & -1.54^{+0.07}_{-0.06} \\
    2.35 & 2.35 & 2.30 & 2.40 & 4317 & -6.28^{+0.09}_{-0.09} & -25.63^{+0.17}_{-0.15} & -3.39^{+0.17}_{-0.19} & -1.64^{+0.06}_{-0.05} \\
    2.45 & 2.45 & 2.40 & 2.50 & 3403 & -6.40^{+0.08}_{-0.07} & -25.86^{+0.14}_{-0.13} & -3.60^{+0.20}_{-0.21} & -1.60^{+0.05}_{-0.05} \\
    2.65 & 2.65 & 2.60 & 2.70 & 1883 & -5.98^{+0.06}_{-0.06} & -25.15^{+0.14}_{-0.13} & -3.13^{+0.12}_{-0.13} & -1.05^{+0.09}_{-0.08} \\
    2.75 & 2.75 & 2.70 & 2.80 & 1135 & -6.30^{+0.08}_{-0.07} & -25.96^{+0.15}_{-0.13} & -3.81^{+0.28}_{-0.31} & -1.34^{+0.07}_{-0.06} \\
    2.85 & 2.85 & 2.80 & 2.90 & 1069 & -6.46^{+0.12}_{-0.10} & -26.23^{+0.23}_{-0.18} & -3.60^{+0.36}_{-0.42} & -1.46^{+0.08}_{-0.07} \\
    2.95 & 2.95 & 2.90 & 3.00 & 1104 & -6.77^{+0.07}_{-0.06} & -26.53^{+0.10}_{-0.09} & -5.05^{+0.60}_{-0.64} & -1.71^{+0.05}_{-0.04} \\
    3.05 & 3.05 & 3.00 & 3.10 & 1127 & -6.77^{+0.08}_{-0.07} & -26.48^{+0.12}_{-0.10} & -4.71^{+0.46}_{-0.52} & -1.70^{+0.06}_{-0.05} \\
    3.15 & 3.15 & 3.10 & 3.20 & 1041 & -7.25^{+0.17}_{-0.12} & -27.09^{+0.25}_{-0.18} & -4.42^{+0.81}_{-1.23} & -1.96^{+0.08}_{-0.06} \\
    3.25 & 3.25 & 3.20 & 3.30 & 815 & -7.33^{+0.13}_{-0.13} & -27.19^{+0.19}_{-0.23} & -4.36^{+0.65}_{-0.76} & -1.93^{+0.06}_{-0.06} \\
    3.34 & 3.35 & 3.30 & 3.40 & 510 & -7.54^{+0.19}_{-0.21} & -27.39^{+0.25}_{-0.37} & -4.90^{+1.36}_{-1.35} & -2.08^{+0.09}_{-0.07} \\
    3.44 & 3.45 & 3.40 & 3.50 & 155 & -6.77^{+0.19}_{-0.20} & -26.64^{+0.42}_{-0.35} & -3.70^{+0.62}_{-0.80} & -1.24^{+0.27}_{-0.22} \\
    3.88 & 3.90 & 3.70 & 4.10 & 628 & -7.92^{+0.12}_{-0.10} & -27.26^{+0.14}_{-0.12} & -4.81^{+0.34}_{-0.39} & -2.07^{+0.10}_{-0.09} \\
    4.35 & 4.40 & 4.10 & 4.70 & 442 & -8.32^{+0.29}_{-0.26} & -27.38^{+0.38}_{-0.31} & -4.19^{+0.41}_{-0.47} & -2.20^{+0.16}_{-0.13} \\
    4.92 & 5.10 & 4.70 & 5.50 & 263 & -9.02^{+0.29}_{-0.22} & -27.88^{+0.36}_{-0.27} & -4.51^{+0.66}_{-0.83} & -2.31^{+0.11}_{-0.08} \\
    6.00 & 6.00 & 5.50 & 6.50 & 66 & -10.60^{+0.71}_{-1.06} & -29.12^{+1.04}_{-1.78} & -4.99^{+0.72}_{-1.22} & -2.40^{+0.10}_{-0.08} \\
    \hline
  \end{tabular}
\end{table*}

@@

%%GW: Not sure what you mean here. I've cut the next two sentences
%As our next step, we combine magnitude bins by binning the quasars
%only in redshifts.  This allows us to estimate the luminosity function
%at each redshift.

In each redshift bin, we model the quasar
luminosity function as a double power law 
\citep[e.g.][]{1988MNRAS.235..935B}
\begin{equation}
  \phi(M) =
  \frac{\phi_*}{10^{0.4(\alpha+1)(M-M_*)}+10^{0.4(\beta+1)(M-M_*)}}
  \label{eqn:dpl}
\end{equation}
with four free parameters: (i) the amplitude $\phi_*$, (ii) the break magnitude $M_*$,
(iii)  the bright-end slope $\alpha$, and (iv) the faint-end slope $\beta$.
%%GW: Maybe give the range of the uniform priors.
By assuming broad, uniform priors, we
obtain posterior probability distributions for these parameters using
MCMC \citep[e.g.,][]{jaynes}.  The joint posterior probability
distribution of the model parameters is then written as
\begin{multline}
  p(\phi_*, M_*, \alpha, \beta | \{M_i, z_i\}) \propto \\ p(\phi_*, M_*,
  \alpha, \beta)p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta),
\end{multline}
where the constant of proportionality is independent of the luminosity
function parameters, and $\{M_i, z_i\}$ denotes the magnitudes and
redshifts of quasars falling in a redshift bin $[z_\mathrm{min},
z_\mathrm{max})$.  We use a uniform prior distribution $p(\phi_*,
M_*, \alpha, \beta)$ and assume that the likelihood
\begin{equation}
  \mathcal{L}\equiv p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta)
\end{equation}
is given by $\phi(M)$ with suitable normalisation.  The negative
logarithm of the likelihood $S\equiv -2\ln\mathcal{L}$ can then be
written as
\begin{multline}
  S = -2\sum_{i=1}^N\ln\phi(M_i, z_i)\\+2\int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, \phi(M,z) f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z}.
  \label{eqn:S}
\end{multline}
where $N$ is the total number of quasars in the redshift bin and the
luminosity integral in the second term on the right hand side is on
the surveyed range of $M$.  We use the \texttt{emcee} code
\citep{2013PASP..125..306F} for MCMC.

@@

The redshift evolution of the four double power law parameters is
shown in Figure~\ref{fig:evoln} and tabulated in Table~\ref{tab:bins}.
We find interesting evolutionary trends in each
of the four parameters.  The break magnitude $M_*$ evolves by more
than eight magnitudes from redshift $z=0$ to $7$.  The amplitude of
the luminosity function $\phi_*$ evolves moderately from $z=0$ to
$z\sim 3$ and then drops by six orders of magnitude to about
$10^{-12}$\,cMpc$^{-3}$mag$^{-1}$ at $z\sim 7$.  The bright end slope $\alpha$ has
significant scatter, but still shows a trend towards more negative
values, i.e., steeper luminosity function bright ends, at high
redshifts.  Finally, the faint end slope of the luminosity function also
shows signs of increasing steepness towards high redshifts.
%%GW: I would not write that beta has a break.
Somewhat similar to the amplitude $\phi_*$, the faint-end slope $\beta$ also
shows a break at $z\sim 3$.  It stays roughly constant up to this
redshift, and the drops to increasingly negative values of close to
$-2.5$ at high redshifts, indicating a steeper faint end.

%%GW: I've commented the following paragraph and replaced it by a sharper version
%Figure~\ref{fig:evoln} reveals further likely systematic errors.
%The behaviour of BOSS quasars, shown by open circles in
%Figure~\ref{fig:evoln} at redshifts $z=2$--$4$ is striking.  In this
%redshift range covered by BOSS the values of the four double power law
%parameters deviate from the general trends.  The faint end of the
%luminosity function of the BOSS quasars is much shallower and the
%break luminosity is fainter by a magnitude compared to the general
%trend expected from a smooth evolution.  The bright end of the
%luminosity function of quasars in this sample is also much flatter
%than that for other samples.  The conspicuousness of the deviation of
%the BOSS sample from the general trends makes it unlikely that the
%luminosity function evolution indicated by BOSS in the redshift range
%$z=2$--$3$ is physical.  Indeed, in this redshift range BOSS quasar
%selection is known to be affected by a colour bias
%\citep{2011ApJ...728...23W}, which can make estimating completeness
%difficult \citep{2006AJ....131.2766R}.

>>>>

The luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.  We evaluate the double
integral in Equation~(\ref{eqn:vi}) by the Euler method, i.e., by
simply summing over the ``tiles'' of the selection function map,
without interpolating between the redshift and luminosity values of
neighbouring tiles. 
%%GW: We should add that interpolation of selection functions is not
%%straightforward (strong gradients) and not unique. Many selection
%%functions are very coarse. V=0 means that the selection function has
%%systematic errors (a QSO was found where the selection probability
%%is zero). By removing them we avoid numerical issues, but
%%underestimate the QLF. Surveys need to report selection functions
%%that are consistent with their actual samples. -- done (GK) 
This may result in $V_i=0$ for some quasars, in which case we remove
them from our analysis.  (Interpolation of selection functions is not
straightforward due to strong gradients.  The presence of objects with
$V_i=0$ implies that the selection function has systematic errors.)
%%GW: Poisson statistics will underestimate the error since
%%uncertainties in the selection function are not considered (all
%%surveys assume a fixed selection function). For small surveys the
%%Poisson error dominates, but for large surveys the error is driven
%%by systematics in the selection function. These facts have been
%%neglected so far, and we should make this point very clearly. -- done (GK)
In each bin, we estimate the uncertainty in the luminosity function by
assuming Poisson statistics \citep{1986ApJ...303..336G} for the number
of quasars, i.e.\ assuming negligible uncertainty in the selection function.
The resultant binned luminosity function estimates are shown by the
points in Figure~\ref{fig:mosaic}.

%%GW: I've sharpened the discussion below. -- done (GK)
As seen in Figure~\ref{fig:mosaic}, the distribution of luminosity
function values in each redshift bin are suggestive of a double power
law form for the QLF.  We will fit such a form below. 
However, in several redshift bins we note a suspicious decline of the
luminosity function at the faint limit of several surveys, for example in
the 2SLAQ sample at $z<2.2$, and in the SDSS sample at $z<1.8$ and $z\sim 4$.
The inconsistency between the SDSS faint end and the deeper
2SLAQ QLF indicates that the SDSS selection function is
systematically overestimated at its magnitude limit.
At $z\sim 4$ the SDSS faint end QLF is inconsistent with the fainter
\citet{2011ApJ...728L..26G} QLF. We identify such discrepant bins
by eye and discard them from our analysis.
%%GW: We should also set the selection functions to zero in these
%%bins. -- done (GK)
The discarded magnitude bins are shown in Figure~\ref{fig:mosaic} by
open circles.

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.7\textwidth]{evolution_individuals.pdf}
    % summary_bins.py 
  \end{center}
  \caption{Redshift evolution of the four double power law parameters from the
    redshift bins shown in Figure~\ref{fig:mosaic}.  Vertical error
    bars show one-sigma (68.26\%) statistical uncertainties, whereas horizontal
    error bars show widths of the redshift bins.  We identify the
    general evolutionary trends of each of these parameters from the
    bins shown by the filled symbols.  The open symbols show bins that
    appear to be offset from these trends, likely due to unknown
    systematic errors.  The open circles at $2.2<z<3.5$ show the
    BOSS sample, while the bins at $z < 0.6$ contain AGN from the SDSS
    and 2SLAQ data sets.  See Section~\ref{sec:bins} for further
    details.}
  \label{fig:evoln}
\end{figure*}

%%GW: I would delete the z_bin column from the table. Otherwise the
%%reader might be confused what you are plotting in the figure. The
%%bin centers and the mean values are very similar anyway.--done (GK)
%%GW: Again 2.5<z<2.6 is missing--done (GK)
%%GW: I would add a column or a comment which redshift bins we
%%consider reliable. I'm afraid lazy people will just take the values
%%without thinking.
\begin{table*}
  % bins_tabulate.py; Nqso added by hand
  \caption{Best-fit double power law luminosity function parameters in
    various redshift bins.  The luminosity function parameters
    $\phi_*$, $M_*$, $\alpha$, and $\beta$ are defined in
    Equation~(\ref{eqn:dpl}), with $\beta$ denoting the faint-end
    slope.  Quasars in each bin have redshifts $z_\mathrm{min}\leq z <
    z_\mathrm{max}$.  The number of quasars in each bin is given by
    $N_\mathrm{qso}$.  The bin centre and sample mean redshift are
    given by $z_\mathrm{bin}$ and $\langle z\rangle$, respectively.
    Errors indicate one-sigma (68.26\%) uncertainties.  These values
    are shown in Figure~\ref{fig:evoln}.  The corresponding luminosity
    functions are shown in Figure~\ref{fig:mosaic}.  See
    Section~\ref{sec:bins} for further details.}
  \label{tab:bins}
  \begin{tabular}{cccr....}
    \hline
    $\langle z\rangle$ &
    $z_\mathrm{min}$ &
    $z_\mathrm{max}$ &
    $N_\mathrm{qso}$ &
    \multicolumn{1}{c}{$\log_{10}(\phi_*/$} &
    \multicolumn{1}{c}{$M_*$} &
    \multicolumn{1}{c}{$\alpha$} &
    \multicolumn{1}{c}{$\beta$} \\

    &
    &
    &
    &
    \multicolumn{1}{c}{cMpc$^{-3}$mag$^{-1}$)} &
    &
    & \\
    \hline
    0.31 & 0.10 & 0.40 & 3632 & -5.62^{+0.09}_{-0.11} & -21.37^{+0.21}_{-0.23} & -2.94^{+0.09}_{-0.10} & -1.22^{+0.18}_{-0.16} \\
    0.50 & 0.40 & 0.60 & 4686 & -6.13^{+0.04}_{-0.04} & -23.00^{+0.08}_{-0.07} & -3.38^{+0.08}_{-0.08} & -1.36^{+0.04}_{-0.04} \\
    0.72 & 0.60 & 0.80 & 4684 & -6.42^{+0.09}_{-0.08} & -24.03^{+0.14}_{-0.13} & -3.56^{+0.12}_{-0.13} & -1.75^{+0.04}_{-0.04} \\
    0.91 & 0.80 & 1.00 & 5248 & -6.39^{+0.08}_{-0.08} & -24.50^{+0.12}_{-0.12} & -3.69^{+0.11}_{-0.11} & -1.73^{+0.06}_{-0.05} \\
    1.10 & 1.00 & 1.20 & 6566 & -6.54^{+0.03}_{-0.03} & -25.15^{+0.04}_{-0.04} & -4.18^{+0.08}_{-0.08} & -1.73^{+0.02}_{-0.02} \\
    1.30 & 1.20 & 1.40 & 7132 & -6.42^{+0.04}_{-0.04} & -25.29^{+0.06}_{-0.05} & -3.97^{+0.08}_{-0.08} & -1.74^{+0.03}_{-0.03} \\
    1.50 & 1.40 & 1.60 & 7771 & -6.51^{+0.03}_{-0.03} & -25.67^{+0.04}_{-0.04} & -4.28^{+0.08}_{-0.08} & -1.76^{+0.02}_{-0.02} \\
    1.71 & 1.60 & 1.80 & 7421 & -6.29^{+0.03}_{-0.03} & -25.56^{+0.05}_{-0.05} & -3.96^{+0.07}_{-0.07} & -1.60^{+0.03}_{-0.03} \\
    1.98 & 1.80 & 2.20 & 10876 & -6.72^{+0.03}_{-0.03} & -26.26^{+0.04}_{-0.04} & -4.21^{+0.07}_{-0.07} & -1.87^{+0.02}_{-0.02} \\
    2.30 & 2.20 & 2.40 & 8419 & -6.16^{+0.07}_{-0.06} & -25.49^{+0.12}_{-0.11} & -3.34^{+0.12}_{-0.12} & -1.61^{+0.04}_{-0.04} \\
    2.45 & 2.40 & 2.50 & 3403 & -6.40^{+0.08}_{-0.07} & -25.86^{+0.14}_{-0.12} & -3.61^{+0.21}_{-0.22} & -1.60^{+0.05}_{-0.04} \\
    2.55 & 2.50 & 2.60 & 2640 & -6.15^{+0.08}_{-0.08} & -25.34^{+0.17}_{-0.16} & -3.32^{+0.15}_{-0.16} & -1.39^{+0.08}_{-0.07} \\
    2.65 & 2.60 & 2.70 & 1883 & -5.98^{+0.05}_{-0.06} & -25.15^{+0.13}_{-0.14} & -3.12^{+0.12}_{-0.12} & -1.05^{+0.09}_{-0.08} \\
    2.75 & 2.70 & 2.80 & 1135 & -6.29^{+0.08}_{-0.07} & -25.94^{+0.15}_{-0.13} & -3.78^{+0.27}_{-0.30} & -1.34^{+0.07}_{-0.06} \\
    2.85 & 2.80 & 2.90 & 1069 & -6.45^{+0.11}_{-0.10} & -26.22^{+0.21}_{-0.18} & -3.61^{+0.35}_{-0.41} & -1.46^{+0.08}_{-0.07} \\
    2.95 & 2.90 & 3.00 & 1104 & -6.76^{+0.07}_{-0.06} & -26.53^{+0.11}_{-0.09} & -5.05^{+0.59}_{-0.63} & -1.71^{+0.05}_{-0.04} \\
    3.05 & 3.00 & 3.10 & 1127 & -6.77^{+0.08}_{-0.07} & -26.48^{+0.12}_{-0.11} & -4.68^{+0.44}_{-0.52} & -1.70^{+0.06}_{-0.05} \\
    3.15 & 3.10 & 3.20 & 1041 & -7.26^{+0.16}_{-0.11} & -27.11^{+0.22}_{-0.17} & -4.38^{+0.76}_{-1.30} & -1.96^{+0.07}_{-0.05} \\
    3.25 & 3.20 & 3.30 & 815 & -7.34^{+0.13}_{-0.13} & -27.20^{+0.20}_{-0.22} & -4.39^{+0.69}_{-0.79} & -1.94^{+0.06}_{-0.05} \\
    3.34 & 3.30 & 3.40 & 510 & -7.54^{+0.21}_{-0.21} & -27.39^{+0.28}_{-0.37} & -4.67^{+1.38}_{-1.44} & -2.08^{+0.09}_{-0.07} \\
    3.44 & 3.40 & 3.50 & 155 & -6.76^{+0.19}_{-0.20} & -26.62^{+0.40}_{-0.36} & -3.67^{+0.62}_{-0.83} & -1.24^{+0.28}_{-0.23} \\
    3.88 & 3.70 & 4.10 & 1204 & -7.91^{+0.11}_{-0.11} & -27.26^{+0.13}_{-0.12} & -4.83^{+0.33}_{-0.38} & -2.07^{+0.10}_{-0.09} \\
    4.35 & 4.10 & 4.70 & 603 & -8.32^{+0.29}_{-0.25} & -27.37^{+0.37}_{-0.30} & -4.20^{+0.41}_{-0.50} & -2.21^{+0.15}_{-0.13} \\
    4.92 & 4.70 & 5.50 & 266 & -9.02^{+0.30}_{-0.21} & -27.89^{+0.37}_{-0.26} & -4.58^{+0.71}_{-0.80} & -2.30^{+0.11}_{-0.08} \\
    6.00 & 5.50 & 6.50 & 66 & -10.66^{+0.74}_{-1.11} & -29.21^{+1.08}_{-1.89} & -5.05^{+0.76}_{-1.18} & -2.41^{+0.10}_{-0.08} \\
    \hline
  \end{tabular}
\end{table*}

@@

%%GW: Not sure what you mean here. I've cut the next two sentences
%As our next step, we combine magnitude bins by binning the quasars
%only in redshifts.  This allows us to estimate the luminosity
%function at each redshift.--done(GK)
In each redshift bin, we model the quasar
luminosity function as a double power law 
\citep[e.g.][]{1988MNRAS.235..935B}
\begin{equation}
  \phi(M) =
  \frac{\phi_*}{10^{0.4(\alpha+1)(M-M_*)}+10^{0.4(\beta+1)(M-M_*)}}
  \label{eqn:dpl}
\end{equation}
with four free parameters: (i) the amplitude $\phi_*$, (ii) the break magnitude $M_*$,
(iii)  the bright-end slope $\alpha$, and (iv) the faint-end slope $\beta$.
%%GW: Maybe give the range of the uniform priors.
By assuming broad, uniform priors, we obtain posterior probability
distributions for these parameters using MCMC \citep[e.g.,][]{jaynes}.
The joint posterior probability distribution of the model parameters
is then written as
\begin{multline}
  p(\phi_*, M_*, \alpha, \beta | \{M_i, z_i\}) \propto \\ p(\phi_*, M_*,
  \alpha, \beta)p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta),
\end{multline}
where the constant of proportionality is independent of the luminosity
function parameters, and $\{M_i, z_i\}$ denotes the magnitudes and
redshifts of quasars falling in a redshift bin $[z_\mathrm{min},
  z_\mathrm{max})$.  We use a uniform prior distribution $p(\phi_*,
  M_*, \alpha, \beta)$ and assume that the likelihood
\begin{equation}
  \mathcal{L}\equiv p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta)
\end{equation}
is given by $\phi(M)$ with suitable normalisation.  The negative
logarithm of the likelihood $S\equiv -2\ln\mathcal{L}$ can then be
written as
\begin{multline}
  S = -2\sum_{i=1}^N\ln\phi(M_i, z_i)\\+2\int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, \phi(M,z) f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z}.
  \label{eqn:S}
\end{multline}
where $N$ is the total number of quasars in the redshift bin and the
luminosity integral in the second term on the right hand side is on
the surveyed range of $M$.  We use the \texttt{emcee} code
\citep{2013PASP..125..306F} for MCMC.

@@

The redshift evolution of the four double power law parameters is
shown in Figure~\ref{fig:evoln} and tabulated in Table~\ref{tab:bins}.
We find interesting evolutionary trends in each of the four
parameters.  The break magnitude $M_*$ evolves by more than eight
magnitudes from redshift $z=0$ to $7$.  The amplitude of the
luminosity function $\phi_*$ evolves moderately from $z=0$ to $z\sim
3$ and then drops by six orders of magnitude to about
$10^{-12}$\,cMpc$^{-3}$mag$^{-1}$ at $z\sim 7$.  The bright end slope
$\alpha$ has significant scatter, but still shows a trend towards more
negative values, i.e., steeper luminosity function bright ends, at
high redshifts.  Finally, the faint end slope of the luminosity
function also shows signs of increasing steepness towards high
redshifts.
%%GW: I would not write that beta has a break.--done(GK)
Somewhat similar to the amplitude $\phi_*$, the faint-end slope
$\beta$ also shows rapid evolution at $z\sim 3$.  It stays roughly
constant up to this redshift, and the drops to increasingly negative
values of close to $-2.5$ at high redshifts, indicating a steeper
faint end.

%%GW: I've commented the following paragraph and replaced it by a
%%sharper version
%Figure~\ref{fig:evoln} reveals further likely systematic errors.
%The behaviour of BOSS quasars, shown by open circles in
%Figure~\ref{fig:evoln} at redshifts $z=2$--$4$ is striking.  In this
%redshift range covered by BOSS the values of the four double power law
%parameters deviate from the general trends.  The faint end of the
%luminosity function of the BOSS quasars is much shallower and the
%break luminosity is fainter by a magnitude compared to the general
%trend expected from a smooth evolution.  The bright end of the
%luminosity function of quasars in this sample is also much flatter
%than that for other samples.  The conspicuousness of the deviation of
%the BOSS sample from the general trends makes it unlikely that the
%luminosity function evolution indicated by BOSS in the redshift range
%$z=2$--$3$ is physical.  Indeed, in this redshift range BOSS quasar
%selection is known to be affected by a colour bias
%\citep{2011ApJ...728...23W}, which can make estimating completeness
%difficult \citep{2006AJ....131.2766R}. --done(GK)
